---
title: <center> Final Project -- Hyperinflation </center>
author: <center> Gaspar Suliansky </center>
date: <center> 11/04/2022 </center>
output: html_document
---
 
<style type = "text/css">
  body{
  font-size: 11 pt;
  font-family: "AppleGothic";
}
</style>

***

```{r global.options, include = FALSE}

knitr::opts_chunk$set(fig.align   = "center")

```


```{r setup, include = FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE)

```

### 1. Load libraries

``` {r message = F}

library(tidyverse)
library(readxl)
library(ggthemes)

# install.packages(c("tidyverse", "readxl", "ggthemes") if necessary

```

***

### 2. Load, tidy and merge data

```{r message = F}

# Latin America (Figures 1 & 2)

argentina <- read_xlsx("/Users/gasparsuliansky/Desktop/Central Banking and Monetary Policy/Final Project/Argentina data.xlsx")
argentina <- add_column(argentina, 
                        country = "Argentina")

peru <- read_xlsx("/Users/gasparsuliansky/Desktop/Central Banking and Monetary Policy/Final Project/Peru data.xlsx")
peru <- add_column(peru, 
                   country = "Peru")

chile <- read_xlsx("/Users/gasparsuliansky/Desktop/Central Banking and Monetary Policy/Final Project/Chile data.xlsx")
chile <- add_column(chile,
                    country = "Chile")

latinamerica_data_1 <- full_join(full_join(argentina, peru), chile) %>% 
  rename(`Total Deficit (% GDP)` = `total deficit`) %>% 
  mutate(`Inflation Rate (log)` = lag(log(`inflation rate`), 1)) %>% 
  pivot_longer(c(5, 18), 
               names_to = "LA_Variables",
               values_to = "LA_Values") %>% 
  subset(select = c(year, 
                    country,
                    LA_Variables,
                    LA_Values))
latinamerica_data_1 # Figure 1

latinamerica_data_2 <- full_join(full_join(argentina, peru), chile) %>% 
  rename(`Total Deficit (% GDP)` = `total deficit`) %>% 
  mutate(`Inflation Rate (log)` = log(`inflation rate`)) %>% 
  subset(select = c(year, 
                    country,
                    `Total Deficit (% GDP)`,
                    `Inflation Rate (log)`))
latinamerica_data_2 # Figure 2

```

```{r message = F}

# Euro Area (Figure 2)

europe_money_supply <- read_csv("/Users/gasparsuliansky/Desktop/Central Banking and Monetary Policy/Final Project/MYAGM2EZM196N.csv")
europe_inflation <- read_csv("/Users/gasparsuliansky/Desktop/Central Banking and Monetary Policy/Final Project/EA19CPALTT01GYM.csv")

europe_data <- inner_join(europe_money_supply,
                          europe_inflation,
                          by = "DATE") %>% 
  rename(`Inflation Rate` = EA19CPALTT01GYM,
         `Money Supply` = MYAGM2EZM196N) %>% 
  mutate(`Monetary Base` = ((`Money Supply` - first(`Money Supply`)) / first(`Money Supply`) + 100),
         `Price Level` = ((`Inflation Rate` - first(`Inflation Rate`)) / first(`Inflation Rate`)) + 100)  %>% 
  pivot_longer(4:5,
               names_to = "EU_Variables",
               values_to = "EU_Values") %>% 
  subset(select = c(DATE,
                     EU_Variables,
                     EU_Values))
europe_data # Figure 3

```

***

### 3. Plot data

&nbsp;

> <font size = "3"> Figure 1 shows time series for two central economic variables: *Total Deficit* (as a % of GDP) and *Inflation Rate* (in a logarithmic scale) -- for each country from 1960 to 2017. The thin black lines showcase the 0% level rate for both variables. Figure 2 shows the same variables in form of a scatter plot to better aknowledge the correlation between the two of them. The dashed lines display the linear regression line for each country. </font>

``` {r warnings = FALSE, message = FALSE, fig.align = 'center'}

# Figure 1

latinamerica_data_1 %>% 
  ggplot(aes(x = year,
             y = LA_Values,
             linetype = LA_Variables,
             col = country)) +
  geom_hline(yintercept = 0,
             col = "black",
             size = 0.20) +
  geom_path() +
  scale_x_continuous(n.breaks = 20) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1,
                                                     scale = 1,
                                                     big.mark = "")) +
  facet_wrap(~ country,
             ncol = 1) +
  guides(col = "none") +
  theme_base() +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        text = element_text(family = "AppleGothic",
                            colour = "black",
                            size = 12),
        axis.text = element_text(colour = "black"),
        plot.caption = element_text(hjust = c(1, 0.5),
                                    margin = margin(t = 15, r = 0, b = 0, l = 0)),
        legend.text = element_text(family = "AppleGothic"),
        legend.background = element_rect(linetype = "solid", 
                                         colour = "black",
                                         size = 0.25,
                                         fill = "grey97"),
        legend.margin = margin(0.10, 0.50, 0.10, 0.10, "cm"),
        strip.text = element_text(size = 11),
        strip.background = element_rect(fill = "grey97"),
        panel.background = element_rect(fill = "grey97"),
        plot.background = element_rect(fill = "grey97"),
        legend.box.background = element_rect(fill = "grey97"),
        legend.key = element_rect(fill = "grey97")) +
  labs(x = "",
       y = "",
       caption = c("Source: Kehoe & Nicolini (2022)",
                   "Figure 1")) +
  scale_color_manual(values = c(Argentina = "skyblue3",
                                Chile = "dodgerblue4",
                                Peru = "firebrick4"))

# Figure 2

latinamerica_data_2 %>% 
  ggplot(aes(x = `Total Deficit (% GDP)`,
             y = `Inflation Rate (log)`,
             fill = country,
             col = country)) +
  geom_jitter(shape = 23,
              alpha = 9 / 10) +
  geom_smooth(method = "lm",
              linetype = "dashed",
              se = F,
              size = 0.5) +
  facet_wrap(~ country,
             ncol = 1) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1,
                                                     scale = 1,
                                                     big.mark = "")) +
  scale_y_continuous(limits = c(0, 10),
                     breaks = c(0, 5, 10), 
                     labels = scales::percent_format(accuracy = 1,
                                                     scale = 1,
                                                     big.mark = "")) +
  theme_base() +
  scale_fill_manual(values = c(Argentina = "skyblue3",
                                Chile = "dodgerblue4",
                                Peru = "firebrick4")) +
  scale_color_manual(values = c(Argentina = "skyblue3",
                               Chile = "dodgerblue4",
                               Peru = "firebrick4")) +
  theme(legend.position = "none",
        text = element_text(family = "AppleGothic",
                            colour = "black",
                            size = 12),
        axis.text = element_text(colour = "black"),
        plot.caption = element_text(hjust = c(1, 0.5),
                                    margin = margin(t = 15, r = 0, b = 0, l = 0)),
        axis.title.x = element_text(margin = margin(t = 15, r = 0, b = 0, l = 0)),
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)),
        strip.text = element_text(size = 11),
        strip.background = element_rect(fill = "grey97"),
        panel.background = element_rect(fill = "grey97"),
        plot.background = element_rect(fill = "grey97")) +
  labs(caption = c("Source: Kehoe & Nicolini (2022)",
                   "Figure 2"))

```

&nbsp;

> <font size = "3"> Figure 3 plots the evolution of the *Monetary Base* and the general *Price Level* for the Euro Area from 1997 to 2015 -- taking the former year as base (1997 = 100). The dashed lines display the linear regression line for each variable. </font>

```{r warnings = FALSE, message = FALSE, fig.align = 'center'}

# Figure 3

europe_data %>%
  ggplot(aes(x = DATE,
             y = EU_Values,
             col = EU_Variables)) +
  geom_line() +
  geom_smooth(method = "lm",
              linetype = "dashed",
              se = F,
              size = 0.5) +
  scale_y_continuous(limits  = c(98, 102)) +
  theme_base() +
  labs(x = "",
       y = "(1997 = 100)",
       title = "Euro Area",
       caption = c("Source: FRED",
                   "Figure 3")) +
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        panel.grid.major = element_blank(),
        text = element_text(family = "AppleGothic",
                            colour = "black",
                            size = 12),
        axis.text = element_text(colour = "black"),
        plot.title = element_text(hjust = 0.5,
                                  margin = margin(t = 10, r = 0, b = 15, l = 0)),
        plot.caption = element_text(hjust = c(1, 0.5),
                                    margin = margin(t = 20, r = 0, b = 0, l = 0)),    
        axis.title.y = element_text(margin = margin(t = 0, r = 15, b = 0, l = 0)),
        panel.background = element_rect(fill = "grey97"),
        plot.background = element_rect(fill = "grey97"),
        legend.box.background = element_rect(fill = "grey97"),
        legend.background = element_rect(linetype = "solid", 
                                         colour = "black",
                                         size = 0.25,
                                         fill = "grey97"),
        legend.key = element_rect(fill = "grey97")) +
  scale_color_manual(values = c("firebrick4",
                                "dodgerblue4"))

```